---
title: "Fakebook"
author: "Adityo_DasGupta"
date: "2023-03-21"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fakebook networking task

The objective of this task is to explore the concept of centrality within networks, specifically in the context of determining the optimal seat or location to occupy during a bus trip from downtown San Francisco to Fakebook. Our ability to connect with other individuals is limited to those within our immediate vicinity and not extend beyond that.


### Visualise the grid with all possible seats
```{r top table}
library(gt)
library(gtExtras)
library(visNetwork)
library(networkD3)
library(ggplot2)
library(dplyr)
library(cluster)
library(tidyverse)
library(igraph)
library(tidygraph)

bus <- data.frame(
  col = rep(c(1:4),times=6), 
  row = rep(c(1:6),each=4),
  passenger=rep('Available',4*6))
  

bus$passenger[bus$row != 5 & bus$col == 1] <- 'NA'
bus$passenger[(bus$row != 5 & bus$col == 4)] <- 'NA'
bus$passenger[(bus$row < 4 & bus$col > 2)] <- 'NA'
bus$passenger[(bus$row == 6 & bus$col > 2)] <- 'NA'


bus$passenger[(bus$row <3 | bus$row >5) & bus$passenger == 'Available' ] <- 'taken'
bus$passenger[bus$col <2 & bus$passenger == 'Available' ] <- 'taken'
bus$passenger[bus$col >2 & bus$passenger == 'Available' & bus$row >4 ] <- 'taken'


ggplot(data=bus,aes(x=row,y=col))+
  geom_point(aes(color = passenger),size=5)+
  labs(title="Grid of Bus no modification")
```

### Visualise the grid with only the seats under contention
```{r top 15 graph}
bus=bus[c(bus$passenger!='NA'),]
bus1=rowid_to_column(bus, "id")

ggplot(data=bus1,aes(x=row,y=col,color=passenger,label=id))+
  geom_point(size=10)+
  geom_text(hjust=.5, vjust=0.5,color="white")+
  labs(title="Grid of Bus with seats under contention")


```

### Get all distances between nodes and keep only which are feasible

```{r nodes}
distances <- daisy(bus1[,c('row','col')], metric = "euclidean")
distances=as.data.frame(as.matrix(distances))


distances=cbind(bus1,distances)

#pivot to make into tabular form
distances <- distances %>% 
  pivot_longer(where(is.numeric) & (!contains(c("id","row","col"))),names_to = "to_seat_id")

#quick change the name of value to distance
distances <- distances %>%
  rename("Distance" = "value")


distances=distances[distances$Distance<=sqrt(2),]

distances%>%
  gt() 
```


### Visualise the network map

```{r edges}
final=distances[,c('id','to_seat_id')]

final=final[final$id!=final$to_seat_id,]
final$to_seat_id=as.numeric(final$to_seat_id)

unique_combinations <- t(apply(final, 1, function(id) sort(id)))
unique_combinations <- unique(unique_combinations)

final_1=data.frame(unique_combinations)

graph <- graph_from_data_frame(final_1, directed=FALSE)

plot(graph, layout=layout.fruchterman.reingold,
     vertex.size = 10,
     vertex.label = V(graph)$name,
     vertex.label.cex = 0.8,
     vertex.label.dist = 1.5,
     vertex.label.color = "black",
     vertex.color = "green")

```


### Get the centralities

```{r network}
betweeness_centrality <- betweenness(graph)


degree_centrality <- degree(graph)


closeness_centrality <- closeness(graph)

output=cbind(data.frame(betweeness_centrality),data.frame(degree_centrality),data.frame(closeness_centrality))
output_1=rowid_to_column(output, "id")
output_1%>%
  gt() 
```


### We observe that the node with ID=7 which is seat 'D' in the original problem is best possible seat


