---
title: "Linkedin_Network"
author: "Adityo_DasGupta"
date: "2023-03-14"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Network Analysis on Linkedin

LinkedIn enables individuals to create valuable professional connections within their industry, highlight their skills and expertise, and explore career prospects. As we delve into my personal network on LinkedIn, we'll examine the companies with whom I have connected the most and uncover potential future opportunities.


### Top 15 Companies in a table
```{r top table}
# Import CSV
connections = read.csv("/Users/adityodasgupta/Downloads/Connections.csv")
attach(connections)


# Create a table with 
connections$name =  paste(connections$First.Name, substr(connections$Last.Name, start = 1, stop = 1), sep = " ")

connections = connections[, c("name", "Company","Position", "Connected.On")]

# create a frequency table
freq_table = table(connections$Company)
freq_table = sort(freq_table, decreasing = TRUE)
top10= head(freq_table, n = 10)


# Display the list as a table
knitr::kable(top10, col.names = c("Company", "Connections"))

```

### Top 10 Companies in a graph
```{r top 15 graph}
# create a bar chart of the frequency table
barplot(top10, main = "Top 10 Connections on Linkedin", 
         ylab = "Number of Connections",
        col = "green", las = 2, cex.names = 0.8)

```

### Creating nodes
I am incorporating the companies my contacts are affiliated with as part of the nodes in my representation.
```{r nodes}
######### Creating nodes

library(tidyverse)

people <- connections %>%
  distinct(name) %>%
  rename(label = name)

companies <- connections %>%
  distinct(Company) %>%
  rename(label = Company)

nodes <- full_join(people, companies, by = "label")
nodes <- rowid_to_column(nodes, "id")
head(nodes)
```


### Creating edges
The edges represent the relationship between the people and the companies.
```{r edges}
#### Creating edges

edges <- connections[, c("name", "Company")]

edges <- edges %>% 
  left_join(nodes, by = c("name" = "label")) %>% 
  rename(from = id)

edges <- edges %>% 
  left_join(nodes, by = c("Company" = "label")) %>% 
  rename(to = id)

edges <- unique(select(edges, from, to))
head(edges)
```


### Graph with network library
It is evident that approximately four companies have the highest number of connections, but the specifics of each are not readily visible.
```{r network}
## Building network
library(network)

routes_network <- network(edges,
                          vertex.attr = nodes,
                          matrix.type = "edgelist",
                          ignore.eval = FALSE)
plot(routes_network, vertex.cex = 1)
```


### Graph with igraph
I utilized igraph to display both the names of companies and individuals, but the sheer volume of connections makes it challenging to discern the details.

```{r igrpah}
## igraph

library(igraph)
routes_igraph <- graph_from_data_frame(d = edges,
                                       vertices = nodes,
                                       directed = TRUE)

plot(routes_igraph,
     vertex.size = 3,
     vertex.label.cex = 0.2,
     edge.arrow.size = 0.01)
```

### Graph with visnetwork library
The implementation of the visnetwork library facilitates the exploration of connections and enables clearer visualization of the companies with the highest number of connections.

```{r visnetwork}
## visNetwork


library(visNetwork)
library(networkD3)
visNetwork(nodes, edges)
```


